---
title: "dataScienceCapstoneProject"
author: "Marcela Castro León"
date: "April 28, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

## Introduction 
Loading libraries

```{r, echo=FALSE, warning=FALSE}
library(NLP)
library(tm)
library(SnowballC)
library(stringi)
library(ggplot2)
library(RColorBrewer)
library(wordcloud)
library(RWeka)
library(slam)
setwd("~/MEGA/MEGAsync/DATA-SCIENCE/work/C10CAPSTONE/dataScienceCaptoneProject/dataScienceCaptoneProject")
```
The original files are data comming from blogs, news and twitter in english language. The files are quite large, 8.9, 10.1 and 23.6 millons of records for blogs, news and twitter respectively. All of them have more than 30 millons of words. In this part we read the files and select randomly 10% of the lines.

```{r, warning=FALSE, echo=FALSE}
readSampleFile<-function(fileName) {
  paste("Reading 20% sample of file ",fileName)
  con=file(fileName,open="r")
  qtotlines<-0     # number of total lines
  qselectedlines <-0   #number of total selected lines
  qtotchar<-0     # number of total chars in file
  qcharselectedlines <-0   #number of total char in selected lines
  finaldata<-NULL
  block=5000
  longmaxline=0;
  while ( TRUE ) {
      line = readLines(con, n = block,skipNul=TRUE)
      if ( length(line) == 0 ) {
          break
      }
      if(sample(1:10,1)==5) {  #random 1:5==5 to select 20% of the lines
          qselectedlines<-qselectedlines+block
          qcharselectedlines<-qcharselectedlines+sum(lengths(strsplit(line, "\\W+")))
          finaldata<-c(finaldata,line)
      }
      lenline=max(nchar(line))
      if(lenline > longmaxline) {
        longmaxline=lenline
      }
      qtotlines <- qtotlines+block
      qtotchar<-qtotchar+sum(lengths(strsplit(line, "\\W+")))
  }
cat(sprintf("File: %s Lines: %.2fM Selected: %.2fM Chars: %.2fM  Selected: %.2fM Max Line %d\n ", fileName, qtotlines/1000000,qselectedlines/1000000,qtotchar/1000000,qcharselectedlines/1000000,longmaxline))
close(con)
return(finaldata)
}

input_sample_file="final_data.txt"
if(!file.exists(input_sample_file)){
    us_blogs <- readSampleFile("final/en_US/en_US.blogs.txt")
    us_news  <- readSampleFile("final/en_US/en_US.news.txt")
    us_twitter  <- readSampleFile("final/en_US/en_US.twitter.txt")
    final_data=c(us_blogs,us_news, us_twitter)
    writeLines(final_data, "final_data.txt")
    } else {
  final_data<-readLines(input_sample_file, encoding = "UTF-8", skipNul=TRUE)
}
```

## Exploring and cleaning final input data 

After reading and selecting 10% of lines of original input data comming from US blogs, news and twitter, we explore the raw lines and chars we have for our analysis.

```{r, echo=TRUE, warning=FALSE}
print("General stats for sample final data")
stri_stats_general(final_data)
```

Using tm package, we create the Corpus object and clean the data by 
*convert to lower case
*remove puntuation  
*remove numbers
*remove extra whitespaces 
*remove common english words

We create a document term matrix to calculate the frequency of each word and a wordclod is plotted.
A sample of 10K lines are choosed in order ti be able to generate the mat

```{r, echo=TRUE, warning=FALSE}
#we select a 000 lines to be able to create the matrix
profanityWords <- read.table("./full-list-of-bad-words-text-file_2018_03_26.txt", header = FALSE)
final_data_sample = sample(final_data,5000, replace = FALSE)
final_data_cleaned <- VCorpus(VectorSource(final_data_sample))
final_data_cleaned <- tm_map(final_data_cleaned, content_transformer(tolower))
final_data_cleaned <- tm_map(final_data_cleaned, content_transformer(tolower))
final_data_cleaned <- tm_map(final_data_cleaned, content_transformer(removePunctuation))
final_data_cleaned <- tm_map(final_data_cleaned, stripWhitespace)
final_data_cleaned <- tm_map(final_data_cleaned, removeWords, stopwords("english"))
#final_data_cleaned <- tm_map(final_data_cleaned, removeWords, unlist(profanityWords))
final_data_cleaned <- tm_map(final_data_cleaned, removeNumbers)
removeURL <- function(x) gsub("http[[:alnum:]]*", "", x)
final_data_cleaned <- tm_map(final_data_cleaned,  content_transformer(removeURL))
BigramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 2, max = 2))
TrigramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 3, max = 3))

dtm<-DocumentTermMatrix(final_data_cleaned)
dtm_bigram <- DocumentTermMatrix(final_data_cleaned, control = list(tokenize = BigramTokenizer, stemming = TRUE))

dtm_trigram <- DocumentTermMatrix(final_data_cleaned, control = list(tokenize = TrigramTokenizer, stemming = TRUE))

dtmmat<-as.matrix(dtm)
freq<-colSums(dtmmat)
freq<-sort(freq,decreasing = TRUE)
words<-names(freq)
wordcloud(words[1:100],freq[1:100])

dtmmat_bigram<-as.matrix(dtm_bigram)
freq_bigram<-colSums(dtmmat_bigram)
freq_bigram<-sort(freq_bigram,decreasing = TRUE)
words_bigram<-names(freq_bigram)
wordcloud(words_bigram[1:100],freq_bigram[1:100])

dtmmat_trigram<-as.matrix(dtm_trigram)
freq_trigram<-colSums(dtmmat_trigram)
freq_trigram<-sort(freq_trigram,decreasing = TRUE)
words_trigram<-names(freq_trigram)
wordcloud(words_trigram[1:100],freq_trigram[1:100])
```

Plot histograms

```{r, echo=TRUE, warning=FALSE}
unigram_freq <- rowapply_simple_triplet_matrix(dtm,sum)
bigram_freq <- rowapply_simple_triplet_matrix(dtm_bigram,sum)
trigram_freq <- rowapply_simple_triplet_matrix(dtm_trigram,sum)
par(mfrow = c(1,3), oma=c(0,0,3,0))
hist(unigram_freq, breaks = 50, main = 'unigram freq', xlab='frequency')
hist(bigram_freq, breaks = 50, main = 'bigram freq', xlab='frequency')
hist(trigram_freq, breaks = 50, main = 'trigram freq', xlab='frequency')
title("NGram Histograms",outer=T)

num <- 30
unigramdf <- head(data.frame(terms=names(freq), freq=freq),n=num) 
bigramdf <- head(data.frame(terms=names(freq_bigram), freq=freq_bigram),n=num)
trigramdf <- head(data.frame(terms=names(freq_trigram), freq=freq_trigram),n=num)

#Plot 1 - Unigram
plot1 <-ggplot(unigramdf,aes(terms, freq))   
plot1 <- plot1 + geom_bar(fill="white", colour=unigramdf$freq,stat="identity") + scale_x_discrete(limits=unigramdf$terms)
plot1 <- plot1 + theme(axis.text.x=element_text(angle=45, hjust=1))  
plot1 <- plot1 + labs(x = "Terms", y="Frequency", title="Top 30 Unigrams in Sample")
plot1  
plot2 <-ggplot(bigramdf,aes(terms, freq))   
plot2 <- plot2 + geom_bar(fill="white", colour=bigramdf$freq,stat="identity") + scale_x_discrete(limits=bigramdf$terms)
plot2 <- plot2 + theme(axis.text.x=element_text(angle=45, hjust=1))  
plot2 <- plot2 + labs(x = "Terms", y="Frequency", title="Top 30 Bigrams in Sample")
plot2
plot3 <-ggplot(trigramdf,aes(terms, freq))   
plot3 <- plot3 + geom_bar(fill="white", colour=trigramdf$freq,stat="identity") + scale_x_discrete(limits=trigramdf$terms)
plot3 <- plot3 + theme(axis.text.x=element_text(angle=45, hjust=1))  
plot3 <- plot3 + labs(x = "Terms", y="Frequency", title="Top 30 Trigrams in Sample")
plot3
```

The histograms are skewed.
We analize how many word most frequency words we need to cover the 50% of total words.

```{r, echo=TRUE, warning=FALSE}
#Words in freq
length(freq)
#sum of frequencies
freqtot<-sum(freq)
freqtot
#sum freq of first 200 words
freq150<-sum(freq[1:150])
#Ratio distribution
150/length(freq)
freq150/freqtot
calcfreq <- function(freq, i) {
    freqtot<-sum(freq)
    freqtot
    freqi<-sum(freq[1:i])
    ratio<-i/length(freq)
    coverage<-freqi/freqtot
    cat(sprintf("Tot words: %d Analized (top frequency) %d words Ratio=%.2f Coverage %.2f\n",length(freq),i,ratio,coverage))
}    
for (i in seq(100,1000,100)) {
  calcfreq(freq,i) 
}


```


