---
title: 'Data Science Captone Quiz 2: Natural language processing I'
author: "Marcela Castro León"
date: "May 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, warning=FALSE, echo=FALSE}
library(NLP)
library(tm)
#read input files
#no input
#returns the union of all the inputs (totlines)
unigram <- readRDS("unigram.RData")
bigram <- readRDS("bigram_nostopwords.RData")
trigram <- readRDS("trigram_nostopwords.RData")
quagram <- readRDS("quagram_nostopwords.RData")

average_frequency<-function(sentence)  {
  sentence_c<-removeWords(sentence,stopwords('en'))
  sentence_c<-stripWhitespace(removeNumbers(removePunctuation(tolower(sentence_c),preserve_intra_word_dashes = TRUE)))
  sentence_split<- strsplit(sentence_c," ")[[1]]
  qwords<-length(sentence_split)
  tot_frequency<-0
  for (i in c(1,2,3,4)) {
    weight_i<-i/10
    last_words<-tail(sentence_split,i)
    #print(last_words)
    if(i==1) {
        freq<-as.integer(head((unigram[unigram$word1==last_words[1],])$freq,1))
    }
    else  
    if(i==2) {
        freq<-as.integer(head((bigram[bigram$word1==last_words[1] 
                                      & bigram$word2 == last_words[2]                               
                                      ,])$freq,1))
        
    } else if(i==3) {
        freq<-as.integer(head((trigram[trigram$word1==last_words[1] 
                                      & trigram$word2 == last_words[2]
                                      & trigram$word3 == last_words[3]
                                      ,])$freq,1))
    }
    else if(i==4) {
        freq<-as.integer(head((quagram[quagram$word1==last_words[1] 
                                      & quagram$word2 == last_words[2]
                                      & quagram$word3 == last_words[3]
                                      & quagram$word4 == last_words[4]
                                      ,])$freq,1))
        
    }
    if(length(freq)==0) freq<-0
    tot_frequency<-tot_frequency+(weight_i*freq)
    #cat(sprintf("with %d words tot_frequency %.2f\n",i, tot_frequency))
  }
  #print(tot_frequency)
  return(tot_frequency)
}
```

## Question 1

For each of the sentence fragments below use your natural language processing algorithm to predict the next word in the sentence.

The guy in front of me just bought a pound of bacon, a bouquet, and a case of

- pretzels

- cheese

- beer

- soda


```{r, warning=FALSE, echo=FALSE}
sentence<-"The guy in front of me just bought a pound of bacon, a bouquet, and a case of"
options<- c('pretzels','cheese', 'beer', 'soda')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```

Answer: beer

## Question 2

You re the reason why I smile everyday. Can you follow me please? It would mean the

world

best

most

universe

```{r, warning=FALSE, echo=FALSE}
sentence<-"You re the reason why I smile everyday. Can you follow me please? It would mean the"
options<- c('world','best', 'most', ' universe')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```

Answer: world

## Question 3
Hey sunshine, can you follow me and make me the

bluest

happiest

smelliest

saddest

```{r, warning=FALSE, echo=FALSE}
sentence<-"Hey sunshine, can you follow me and make me the"
options<- c('bluest','happiest', 'smelliest', 'saddest')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```


Answer: happiest


## Question 4

Very early observations on the Bills game: Offense still struggling but the

players

defense

referees

crowd

```{r, warning=FALSE, echo=FALSE}
sentence<-"Very early observations on the Bills game: Offense still struggling but the"
options<- c('players','defense', 'referees', 'crowd')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```
Answer: defense

## Question 5

Go on a romantic date at the

movies

mall

grocery

beach

```{r, warning=FALSE, echo=FALSE}
sentence<-"Go on a romantic date at the"
options<- c('movies','mall', 'grocery', 'beach')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```
Answer: beach

## Question 6

Well I'm pretty sure my granny has some old bagpipes in her garage I'll dust them off and be on my

motorcycle

way

horse

phone

```{r, warning=FALSE, echo=FALSE}
sentence<-"Well I'm pretty sure my granny has some old bagpipes in her garage I'll dust them off and be on my"
options<-c('motorcycle','way', 'horse', 'phone')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```
Answer: way

## Question 7
Ohhhhh #PointBreak is on tomorrow. Love that film and haven't seen it in quite some

time

thing

years

weeks


```{r, warning=FALSE, echo=FALSE}
sentence<-"Ohhhhh #PointBreak is on tomorrow. Love that film and haven't seen it in quite some"
options<-c('time','thing', 'years', 'weeks')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```

Answer: time

## Question 8

After the ice bucket challenge Louis will push his long wet hair out of his eyes with his little

ears

eyes

toes

fingers

```{r, warning=FALSE, echo=FALSE}
sentence<-"After the ice bucket challenge Louis will push his long wet hair out of his eyes with his little"
options<-c('ears','eyes', 'toes', 'fingers')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```
Answer: fingers

## Question 9

Be grateful for the good times and keep the faith during the

worse

bad

hard

sad

```{r, warning=FALSE, echo=FALSE}
sentence<-"Be grateful for the good times and keep the faith during the"
options<-c('worse','bad', 'hard', 'sad')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```
Answer:bad

## Question 10
If this isn't the cutest thing you've ever seen, then you must be

asleep

callous

insane

insensitive


```{r, warning=FALSE, echo=FALSE}
sentence<-"If this isn't the cutest thing you've ever seen, then you must be"
options<-c('asleep','callous', 'insane', 'insensitive')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```

Answer: asleep