---
title: "Data Science Capstone Project Quiz 3"
author: "Marcela Castro León"
date: "6 de mayo de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

## Preparation 

Read data, create corpus using a sampling size and save unigrams, bigrams, trigrams and quagrams 


```{r, warning=FALSE, echo=FALSE}
#read input files
#no input
#returns the union of all the inputs (totlines)
unigram <- readRDS("unigram.RData")
bigram <- readRDS("bigram_nostopwords.RData")
trigram <- readRDS("trigram_nostopwords.RData")
quagram <- readRDS("quagram_nostopwords.RData")

average_frequency<-function(sentence)  {
  sentence_c<-removeWords(sentence,stopwords('en'))
  sentence_c<-stripWhitespace(removeNumbers(removePunctuation(tolower(sentence_c),preserve_intra_word_dashes = TRUE)))
  sentence_split<- strsplit(sentence_c," ")[[1]]
  qwords<-length(sentence_split)
  tot_frequency<-0
  for (i in c(1,2,3,4)) {
    weight_i<-i/10
    last_words<-tail(sentence_split,i)
    #print(last_words)
    if(i==1) {
        freq<-as.integer(head((unigram[unigram$word1==last_words[1],])$freq,1))
    }
    else  
    if(i==2) {
        freq<-as.integer(head((bigram[bigram$word1==last_words[1] 
                                      & bigram$word2 == last_words[2]                               
                                      ,])$freq,1))
        
    } else if(i==3) {
        freq<-as.integer(head((trigram[trigram$word1==last_words[1] 
                                      & trigram$word2 == last_words[2]
                                      & trigram$word3 == last_words[3]
                                      ,])$freq,1))
    }
    else if(i==4) {
        freq<-as.integer(head((quagram[quagram$word1==last_words[1] 
                                      & quagram$word2 == last_words[2]
                                      & quagram$word3 == last_words[3]
                                      & quagram$word4 == last_words[4]
                                      ,])$freq,1))
        
    }
    if(length(freq)==0) freq<-0
    tot_frequency<-tot_frequency+(weight_i*freq)
    #cat(sprintf("with %d words tot_frequency %.2f\n",i, tot_frequency))
  }
  #print(tot_frequency)
  return(tot_frequency)
}
```

## Question 1

For each of the sentence fragments below use your natural language processing algorithm to predict the next word in the sentence.

When you breathe, I want to be the air for you. I'll be there for you, I'd live and I'd

give

eat

die

sleep

```{r, warning=FALSE, echo=FALSE}
sentence<-"When you breathe, I want to be the air for you. I'll be there for you, I'd live and I'd "
options<- c('give','eat','die','sleep')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```

Answer: die

## Question 2
Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his

marital

spiritual

horticultural

financial

```{r, warning=FALSE, echo=FALSE}
sentence<-"Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his "
options<- c('marital','spiritual','horticultural','financial')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```
Answer marital

## Question 3

I'd give anything to see arctic monkeys this

month

decade

weekend

morning


```{r, warning=FALSE, echo=FALSE}
sentence<-"I'd give anything to see arctic monkeys this "
options<- c('month','decade','weekend','morning')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```

Answer: weekend


## Question 4

Talking to your mom has the same effect as a hug and helps reduce your

stress

sleepiness

hunger

happiness

```{r, warning=FALSE, echo=FALSE}
sentence<-"Talking to your mom has the same effect as a hug and helps reduce your "
options<- c('stress','sleepiness','hunger','happiness')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```

Answer stress

## Question 5

When you were in Holland you were like 1 inch away from me but you hadn't time to take a

walk

look

minute

picture

```{r, warning=FALSE, echo=FALSE}
sentence<-"Talking to your mom has the same effect as a hug and helps reduce your "
options<- c('walk','look','minute','picture')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i])
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```

Answer: picture

## Question 6

I'd just like all of these questions answered, a presentation of evidence, and a jury to settle the

incident

case

matter

account

```{r, warning=FALSE, echo=FALSE}
sentence<-"I'd just like all of these questions answered, a presentation of evidence, and a jury to settle the"
options<- c('incident','case','matter','account')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i]," ")
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```
Answer: matter


## Question 7

I can't deal with unsymetrical things. I can't even hold an uneven number of bags of groceries in each

finger

arm

toe

hand

```{r, warning=FALSE, echo=FALSE}
sentence<-"I can't deal with unsymetrical things. I can't even hold an uneven number of bags of groceries in each"
options<- c('finger','arm','toe','hand')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i]," ")
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```
Answer: hand

## Question 8
Every inch of you is perfect from the bottom to the

middle

side

center

top

```{r, warning=FALSE, echo=FALSE}
sentence<-"Every inch of you is perfect from the bottom to the"
options<- c('middle','side','center','top')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i]," ")
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```

Answer: top

## Question 9

I’m thankful my childhood was filled with imagination and bruises from playing

outside

daily

inside

weekly

```{r, warning=FALSE, echo=FALSE}
sentence<-"I’m thankful my childhood was filled with imagination and bruises from playing"
options<- c('outside','daily','inside','weekly')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i]," ")
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```

Answer Outside


## Question 10

I like how the same people are in almost all of Adam Sandler's

stories

pictures

novels

movies

```{r, warning=FALSE, echo=FALSE}
sentence<-"I like how the same people are in almost all of Adam Sandler's"
options<- c('stories','pictures','novels','movies')
for(i in 1:length(options)) {
  sentence_n<-paste(sentence, options[i]," ")
  print(sentence_n)
  a<-average_frequency(sentence_n)
  print(a)
}
```

Answer: movies
