---
title: 'Data Science Captone Quiz 2: Natural language processing I'
author: "Marcela Castro León"
date: "May 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, echo=FALSE}
suppressMessages(library(ggplot2))
suppressMessages(library(NLP))
library(tm)
library(SnowballC)
library(stringi)
library(RColorBrewer)
library(wordcloud)
library(RWeka)
library(slam)
```

```{r, warning=FALSE, echo=FALSE}
filterVchar<-function(totlines,stringToMatch) {
  match<-totlines[grepl(stringToMatch,totlines)]
  cat(sprintf("Lines: %d with string %s Lines=%d\n ",length(totlines), stringToMatch, length(match)))
return(match)
}

getFinalMatch2<-function(stringToMatch) {
  #in_blogs <- readLines("final/en_US/en_US.blogs.txt",skipNul=TRUE)
  #in_news  <- readLines("final/en_US/en_US.news.txt",skipNul=TRUE)
  #in_twitter  <- readLines("final/en_US/en_US.twitter.txt",skipNul=TRUE)
  #totlines<-c(in_blogs,in_news, in_twitter)
  final_match=filterVchar(totlines,stringToMatch)
  print("General stats for final match")
  stri_stats_general(final_match)
  return(final_match)
}

readFilterFile<-function(fileName,stringToMatch) {
  cat(sprintf("Selecting lines with %s in %s\n",stringToMatch, fileName))
  totlines<-readLines(fileName,skipNul=TRUE)
  match<-totlines[grepl(stringToMatch,totlines)]
  cat(sprintf("File: %s Lines: %d with string %s Lines=%d\n ",  basename(fileName), length(totlines), stringToMatch, length(match)))
return(match)
}

getFinalMatch<-function(stringToMatch) {
  in_blogs <- readFilterFile("final/en_US/en_US.blogs.txt",stringToMatch)
  in_news  <- readFilterFile("final/en_US/en_US.news.txt",stringToMatch)
  in_twitter  <- readFilterFile("final/en_US/en_US.twitter.txt",stringToMatch)
  final_match=c(in_blogs,in_news, in_twitter)
  print("General stats for final match")
  stri_stats_general(final_match)
  return(final_match)
}

getCleanedCorpus<-function(final_match) {
  set.seed(1234)
  final_data_sample<-final_match
  #creating corpus object to use tn functions
  final_data_cleaned <- VCorpus(VectorSource(final_data_sample))
  #cleaning operations on sample text: lower, remove punctuation, remove common words, etc
  final_data_cleaned <- tm_map(final_data_cleaned, content_transformer(tolower))
  final_data_cleaned <- tm_map(final_data_cleaned, content_transformer(removePunctuation))
  final_data_cleaned <- tm_map(final_data_cleaned, stripWhitespace)
  final_data_cleaned <- tm_map(final_data_cleaned, removeWords, stopwords("english"))
  final_data_cleaned <- tm_map(final_data_cleaned, removeNumbers)
  removeURL <- function(x) gsub("http[[:alnum:]]*", "", x)
  final_data_cleaned <- tm_map(final_data_cleaned,  content_transformer(removeURL))
  return(final_data_cleaned)
}

getFreqVector<-function(final_data_cleaned, num_grams) {
  BigramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 2, max = 2))
  TrigramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 3, max = 3))
  QuagramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 4, max = 4))
  QuigramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 5, max = 5))
  if(num_grams==2)  {funtoken<-BigramTokenizer}
  else if(num_grams==3) {funtoken<-TrigramTokenizer}
  else if(num_grams==4) {funtoken<-QuagramTokenizer}
  else {funtoken<-QuigramTokenizer}
  dtm <- DocumentTermMatrix(final_data_cleaned, control = list(tokenize = funtoken, stemming = TRUE))  
  dtmmat<-as.matrix(dtm)
  freq<-colSums(dtmmat)
  freq<-sort(freq,decreasing = TRUE)
  return(freq)
}
```



## Question 1

For each of the sentence fragments below use your natural language processing algorithm to predict the next word in the sentence.

The guy in front of me just bought a pound of bacon, a bouquet, and a case of

- pretzels

- cheese

- beer

- soda


```{r, echo=TRUE, warning=FALSE}
stringToMatch="([Aa]+ +[Cc]ase+ +[Oo]f+ +[^ ]+ )"
test<-c('pretzels','cheese', 'beer', 'soda')
final_match<-getFinalMatch2(stringToMatch)
final_data_cleaned<-getCleanedCorpus(final_match)
freq_trigram<-getFreqVector(final_data_cleaned, 3)  
for (i in 1:length(test)) {
  matches<-freq_trigram[grepl(test[i], names(freq_trigram))]
  print(paste("Matches for ",test[i]))
  print(matches)
}
```

Answer: beer

## Question 2

You re the reason why I smile everyday. Can you follow me please? It would mean the

world

best

most

universe

```{r, echo=TRUE, warning=FALSE}
stringToMatch="([Ww]ould+ +[Mm]ean+ +[Tt]he+ +[^ ]+ )"
test<-c('world','best', 'most', 'universe')
final_match<-getFinalMatch(stringToMatch)
final_data_cleaned<-getCleanedCorpus(final_match)
freq_trigram<-getFreqVector(final_data_cleaned, 3)  
for (i in 1:length(test)) {
  matches<-freq_trigram[grepl(test[i], names(freq_trigram))]
  print(paste("Matches for ",test[i]))
  print(matches)
}
```

Answer: world

## Question 3
Hey sunshine, can you follow me and make me the

bluest

happiest

smelliest

saddest

```{r, echo=TRUE, warning=FALSE}
stringToMatch="([Mm]ake+ +[Mm]e+ +[Tt]he+ +[^ ]+ )"
test<-c('bluest','happiest', 'smelliest', 'saddest')
final_match<-getFinalMatch2(stringToMatch)
final_data_cleaned<-getCleanedCorpus(final_match)
freq_trigram<-getFreqVector(final_data_cleaned, 4)  
for (i in 1:length(test)) {
  matches<-freq_trigram[grepl(test[i], names(freq_trigram))]
  print(paste("Matches for ",test[i]))
  print(matches)
}
```
Answer: happiest


## Question 4

Very early observations on the Bills game: Offense still struggling but the

players

defense

referees

crowd

```{r, echo=TRUE, warning=FALSE}
stringToMatch="([Ss]truggling+ +[^ ])" 
test<-c('players','defense', 'referees', 'crowd')
final_match<-getFinalMatch2(stringToMatch)
final_data_cleaned<-getCleanedCorpus(final_match)
freq_trigram<-getFreqVector(final_data_cleaned, 3)  
for (i in 1:length(test)) {
  matches<-freq_trigram[grepl(test[i], names(freq_trigram))]
  print(paste("Matches for ",test[i]))
  print(matches)
}
```

Answer: defense

## Question 5

Go on a romantic date at the

movies

mall

grocery

beach

```{r, echo=TRUE, warning=FALSE}
stringToMatch="([Dd]ate+ +[Aa]t+ +[^ ]+ )"
test<-c('movies','mall', 'grocery', 'beach')
final_match<-getFinalMatch2(stringToMatch)
final_data_cleaned<-getCleanedCorpus(final_match)
freq_trigram<-getFreqVector(final_data_cleaned, 4)  
for (i in 1:length(test)) {
  matches<-freq_trigram[grepl(test[i], names(freq_trigram))]
  print(paste("Matches for ",test[i]))
  print(matches)
}
```

Answer:

## Question 6

Well I'm pretty sure my granny has some old bagpipes in her garage I'll dust them off and be on my

motorcycle

way

horse

phone

```{r, echo=TRUE, warning=FALSE}
stringToMatch="([Bb]e+ +[Oo]n+ +[Mm]y+ +[^ ]+ )"
test<-c('motorcycle','way', 'horse', 'phone')
final_match<-getFinalMatch2(stringToMatch)
final_data_cleaned<-getCleanedCorpus(final_match)
freq_trigram<-getFreqVector(final_data_cleaned, 4)  
for (i in 1:length(test)) {
  matches<-freq_trigram[grepl(test[i], names(freq_trigram))]
  print(paste("Matches for ",test[i]))
  print(matches)
}
```
Answer: way

## Question 7
Ohhhhh #PointBreak is on tomorrow. Love that film and haven't seen it in quite some

time

thing

years

weeks

```{r, echo=TRUE, warning=FALSE}
stringToMatch="([Qq]uite+ +[Ss]ome+ +[^ ]+ )"
test<-c('time','thing', 'years', 'weeks')
final_match<-getFinalMatch2(stringToMatch)
final_data_cleaned<-getCleanedCorpus(final_match)
freq_trigram<-getFreqVector(final_data_cleaned, 4)  
for (i in 1:length(test)) {
  matches<-freq_trigram[grepl(test[i], names(freq_trigram))]
  print(paste("Matches for ",test[i]))
  print(matches)
}
```
Answer: time

## Question 8

After the ice bucket challenge Louis will push his long wet hair out of his eyes with his little

ears

eyes

toes

fingers

```{r, echo=TRUE, warning=FALSE}
stringToMatch="([Hh]is+ +[Ll]ittle+ +[^ ]+ )" 
test<-c('little ears','little eyes', 'little toes', 'little fingers')
final_match<-getFinalMatch2(stringToMatch)
final_data_cleaned<-getCleanedCorpus(final_match)
freq_trigram<-getFreqVector(final_data_cleaned, 3)  
for (i in 1:length(test)) {
  matches<-freq_trigram[grepl(test[i], names(freq_trigram))]
  print(paste("Matches for ",test[i]))
  print(matches)
}
```
Answer: fingers

## Question 9

Be grateful for the good times and keep the faith during the

worse

bad

hard

sad

```{r, echo=TRUE, warning=FALSE}
stringToMatch="[Dd]uring the bad"
test<-c('worse','bad', 'hard', 'sad')
final_match<-getFinalMatch2(stringToMatch)
final_data_cleaned<-getCleanedCorpus(final_match)
freq_trigram<-getFreqVector(final_data_cleaned, 2)  
for (i in 1:length(test)) {
  matches<-freq_trigram[grepl(test[i], names(freq_trigram))]
  print(paste("Matches for ",test[i]))
  print(matches)
}
```
Answer:bad

## Question 10
If this isn't the cutest thing you've ever seen, then you must be

asleep

callous

insane

insensitive

```{r, echo=TRUE, warning=FALSE}
stringToMatch="([Mm]ust+ +[Bb]e+ +[^ ]+ )"
test<-c('asleep','callous', 'insane', 'insensitive')
final_match<-getFinalMatch2(stringToMatch)
final_data_cleaned<-getCleanedCorpus(final_match)
freq_trigram<-getFreqVector(final_data_cleaned, 2)  
for (i in 1:length(test)) {
  matches<-freq_trigram[grepl(test[i], names(freq_trigram))]
  print(paste("Matches for ",test[i]))
  print(matches)
}
```

Answer: